---
layout: page
title: ""
date: 2018-01-20
author: "Thomas M. Massie"
output:
  html_document:
    keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include = FALSE}
rm(list=ls())
```


```{r include = FALSE}
# Load libraries.
library(repmis)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(lubridate)
library(scales)
library(chron)
library(ggmap)
library(sp)
library(leaflet)
library(maps)
library(maptools)
library(cowplot)
library(RColorBrewer)
library(ggpubr)
library(showtext)
library(broom)
library(shiny)
```



# Storm data exploration (3/n)



## Loading the environment

Loading the workspace that has been saved in the first part of the data exploration.
```{r message = FALSE}
# Entire workspace.
source_data("https://github.com/thomassie/Storms/blob/master/Exploration/StormDataWorkSpace.RData?raw=true")
```


When do the storms typicall occur during a year?
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.pressure.year.main <- ggplot(data = dd,
                                  aes(x = DateTimeSameYear,
                                      y = Pressure)) + 
  geom_point(aes(colour = Ocean),
             size = 0.8,
             alpha = 0.1) +
  # geom_smooth(method = "lm",
  #             color = "#333333") +
  labs(x = "Month", 
       y = "Minimum central pressure (mbar)",
       title = "When do hurricans and typhoons occur?",
       subtitle = expression("Annual minima in central pressure records."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = -0.5, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0, vjust = -5, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1.6, vjust = -0.1, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.title = element_blank(),
        legend.justification=c(0,1),
        legend.position=c(1.02, 0.3),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(size = 10)) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  # rremove("legend") +
  scale_x_datetime(labels = date_format("%b"),
                   date_breaks = "month",
                   limits = c(min(dd$DateTimeSameYear), max(dd$DateTimeSameYear))) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

plot.pressure.year.dens.x <- axis_canvas(plot.pressure.year.main, axis = "x") + 
  geom_density(data = dd, aes(x = as.numeric(DateTimeSameYear), fill = Ocean),
               alpha = 0.6, size = 0.2) +
  scale_fill_manual(values = c("#FF281E", "#0090CF"))

# ...and one on the right site.
plot.pressure.year.dens.y <- axis_canvas(plot.pressure.year.main, axis = "y", coord_flip = TRUE) +
  geom_density(data = dd, aes(x = as.numeric(Pressure), fill = Ocean),
               alpha = 0.6, size = 0.2) +
  scale_fill_manual(values = c("#FF281E", "#0090CF")) +
  coord_flip()
```

```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.pressure.year.1 <- insert_xaxis_grob(plot.pressure.year.main, 
                                          plot.pressure.year.dens.x, 
                                          grid::unit(0.2, "null"), 
                                          position = "top")
plot.pressure.year.2 <- insert_yaxis_grob(plot.pressure.year.1, 
                                          plot.pressure.year.dens.y, 
                                          grid::unit(.2, "null"), 
                                          position = "right")
ggdraw(plot.pressure.year.2)
```


When do the storms typicall occur during a year?
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE, echo = FALSE}
plot.duration.main <- ggplot(data = dd,
                             aes(x = Duration,
                                 y = Pressure)) + 
  geom_vline(xintercept = 0, size = 0.8, colour = "#3C3C3C") +
  geom_point(aes(colour = Ocean),
             size = 0.8,
             alpha = 0.1) +
  labs(x = "Duration (days)", 
       y = "Minimum central pressure (mbar)",
       title = expression("When do storms reach their maximum intensity?"),
       subtitle = expression("Most storms showed their lowest minimum in \ncentral pressure briefly after they built up."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = -0.5, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0.9, vjust = -7, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1.6, vjust = -0.1, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.title = element_blank(),
        legend.justification=c(0,1),
        legend.position=c(1.02, 0.3),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(size = 10)) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

# A density plot on top of the main plot.
# TREAT the date.time AS NUMERIC!!!
plot.duration.dens.x <- axis_canvas(plot.duration.main, axis = "x") + 
  geom_density(data = dd, aes(x = Duration, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  scale_fill_manual(values = c("#FF281E", "#0090CF"))

# ...and one on the right site.
plot.duration.dens.y <- axis_canvas(plot.duration.main, axis = "y", coord_flip = TRUE) +
  geom_density(data = dd, aes(x = Pressure, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  scale_fill_manual(values = c("#FF281E", "#0090CF")) +
  coord_flip()
```

```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE, echo = FALSE}
plot.duration.1 <- insert_xaxis_grob(plot.duration.main, 
                                     plot.duration.dens.x, 
                                     grid::unit(0.2, "null"), 
                                     position = "top")
plot.duration.2 <- insert_yaxis_grob(plot.duration.1, 
                                     plot.duration.dens.y, 
                                     grid::unit(.2, "null"), 
                                     position = "right")
ggdraw(plot.duration.2)
```


After what time do storms typically reach their maximum intensity?
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE, echo = FALSE}
plot.duration.storm.main <- ggplot(data = dd.sum,
                                   aes(x = MinPressureAfter,
                                       y = MinPressure)) + 
  geom_vline(xintercept = 0, size = 0.8, colour = "#3C3C3C") +
  geom_point(aes(colour = Ocean,
                 size = MaxWindKPH),
             show.legend = TRUE,
             # size = 0.8,
             alpha = 0.2) +
  labs(x = "Duration at lowest minimum pressure (days)", 
       y = "Minimum central pressure (mbar)",
       title = expression("When do storms reach their maximum intensity?"),
       subtitle = expression("Most storms show their lowest minimum \npressure briefly after they built up."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = -0.5, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0.9, vjust = -7, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1.6, vjust = -0.1, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.justification=c(0,1),
        legend.position=c(1.02, 0.3),
        legend.background = element_blank(),
        legend.title=element_text(size=10),
        legend.text = element_text(size = 10)) +
  labs(colour = "Ocean",
       size = expression("Wind speed \n(km/h)")) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

# A density plot on top of the main plot.
# TREAT the date.time AS NUMERIC!!!
plot.duration.storm.dens.x <- axis_canvas(plot.duration.storm.main, axis = "x") + 
  geom_density(data = dd.sum, aes(x = MinPressureAfter, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  scale_fill_manual(values = c("#FF281E", "#0090CF"))

# ...and one on the right site.
plot.duration.storm.dens.y <- axis_canvas(plot.duration.storm.main, axis = "y", coord_flip = TRUE) +
  geom_density(data = dd.sum, aes(x = MinPressure, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  scale_fill_manual(values = c("#FF281E", "#0090CF")) +
  coord_flip()
```

```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE, echo = FALSE}
plot.duration.storm.1 <- insert_xaxis_grob(plot.duration.storm.main, 
                                           plot.duration.storm.dens.x, 
                                           grid::unit(0.2, "null"), 
                                           position = "top")
plot.duration.storm.2 <- insert_yaxis_grob(plot.duration.storm.1, 
                                           plot.duration.storm.dens.y, 
                                           grid::unit(.2, "null"), 
                                           position = "right")
ggdraw(plot.duration.storm.2)
```


However, the time at which a storm reaches its maximum intensity has to be related to its total duration.
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE, echo = FALSE}
plot.duration.storm.main <- ggplot(data = dd.sum,
                                   aes(x = MinPressureAfter/MaxDuration *100,
                                       y = MinPressure)) + 
  geom_vline(xintercept = 0, size = 0.8, colour = "#3C3C3C") +
  geom_point(aes(colour = Ocean,
                 size = MaxWindKPH),
             # size = MaxWindKPH^4/100000),
             show.legend = TRUE,
             # size = 0.8,
             alpha = 0.2) +
  # geom_smooth(method = "lm",
  #             color = "#333333") +
  labs(x = "Duration at lowest minimum pressure in relation to total duration (%)", 
       y = "Minimum central pressure (mbar)",
       title = expression("When do storms reach their maximum intensity?"),
       subtitle = expression("Most storms are at their maximum intensity at around half-time."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = 0, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0, vjust = 0.5, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1.6, vjust = -0.1, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.justification=c(0,1),
        # legend.title = element_blank(),
        legend.position=c(1.02, 0.3),
        legend.background = element_blank(),
        legend.title=element_text(size=10),
        # legend.key = element_blank(),
        legend.text = element_text(size = 10)) +
  labs(colour = "Ocean",
       size = expression("Wind speed \n(km/h)")) +
  # geom_text(data = dd.lines, aes(label = year(int), x = int, y = -Inf), angle = 0, inherit.aes = F, hjust = -0.2, vjust = -36, size = 3.5) +
  # ggpubr::color_palette("jco") +
  # scale_colour_manual(values = c("#B78A3F", "#58758C")) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  # rremove("legend") +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

# A density plot on top of the main plot.
# TREAT the date.time AS NUMERIC!!!
plot.duration.storm.dens.x <- axis_canvas(plot.duration.storm.main, axis = "x") + 
  geom_density(data = dd.sum, aes(x = MinPressureAfter/MaxDuration *100, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # geom_histogram(data = dd.sum, aes(x = MinPressureAfter/MaxDuration *100, fill = Ocean),
  #                alpha = 0.6, size = 0.2,
  #                binwidth = 2) +
  # ggpubr::color_palette("jco")
  # scale_fill_manual(values = c("#B78A3F", "#58758C"))
  scale_fill_manual(values = c("#FF281E", "#0090CF"))

# ...and one on the right site.
plot.duration.storm.dens.y <- axis_canvas(plot.duration.storm.main, axis = "y", coord_flip = TRUE) +
  geom_density(data = dd.sum, aes(x = MinPressure, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # geom_histogram(data = dd.sum, aes(x = MinPressure, fill = Ocean),
  #                alpha = 0.6, size = 0.2,
  #                binwidth = 2) +
  # ggpubr::color_palette("jco") +
  # scale_fill_manual(values = c("#B78A3F", "#58758C")) +
  scale_fill_manual(values = c("#FF281E", "#0090CF")) +
  coord_flip()
```

```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE, echo = FALSE}
plot.duration.storm.1 <- insert_xaxis_grob(plot.duration.storm.main, 
                                           plot.duration.storm.dens.x, 
                                           grid::unit(0.2, "null"), 
                                           position = "top")
plot.duration.storm.2 <- insert_yaxis_grob(plot.duration.storm.1, 
                                           plot.duration.storm.dens.y, 
                                           grid::unit(.2, "null"), 
                                           position = "right")
ggdraw(plot.duration.storm.2)
```


Next, one can ask for the regions that are most storm prone.
```{r}
plot.pressure.main <- ggplot(data = dd,
                             aes(x = Lon -360,
                                 y = Lat)) +
  stat_density_2d(geom = "raster",
                  aes(fill = ..density..),
                  contour = FALSE,
                  alpha = 0.9,
                  show.legend = FALSE)
# scale_fill_gradientn(colours = rev( brewer.pal( 9, "Blues" ))) +
```







