---
layout: page
title: ""
date: 2018-01-14
author: "Thomas M. Massie"
output:
  html_document:
    keep_md: true
# runtime: shiny  # to make sure leaflet objects are rendered on site
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include = FALSE}
rm(list=ls())
```


```{r include = FALSE}
# Load libraries.
library(repmis)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(lubridate)
library(scales)
library(chron)
library(ggmap)
library(sp)
library(leaflet)
library(maps)
library(maptools)
library(cowplot)
library(RColorBrewer)
library(ggpubr)
library(showtext)
library(broom)
library(shiny)
```



# Storm data exploration (2/n)



## Loading the environment

Loading the workspace that has been saved in the first part of the data exploration.
```{r message = FALSE}
# Entire workspace.
source_data("https://github.com/thomassie/Storms/blob/master/Exploration/StormDataWorkSpace.RData?raw=true")
```


This time I will have a look at specific parts of the data set according to the choices I made before (i.e., a defined time interval; see first part of data exploration).
```{r warning = FALSE, echo = FALSE, include = FALSE}
# All storms between 'year.min' and 'year.max'. 
dd.select <- dd %>% 
  filter(year.min <= year(DateTime) & year(DateTime) <= year.max)

# The summary statistics for the selected data.
dd.select.sum <- dd.sum %>%
  filter(Key %in% dd.select$Key)

# Longest duration (selection).
topn.select.Duration      <- arrange(dd.select.sum, desc(dd.select.sum$MaxDuration))[1:n.select,]
# Strongest wind recorded (selection)
topn.select.Strength.Max  <- arrange(dd.select.sum, desc(dd.select.sum$MaxWindKPH))[1:n.select,]
# Lowest pressure (selection)
topn.select.Lowest.       <- arrange(dd.select.sum, -desc(dd.select.sum$MaxWindKPH))[1:n.select,]
# Minimum pressure (selection)
topn.select.Pressure.Min  <- arrange(dd.select.sum, -desc(dd.select.sum$MinPressure))[1:n.select,]
```


Let's have a look at these storms. First, I want to see which are the storms that lasted the longest and where they ocurred. (I use a data set called 'dd.s' indicating a selection of the entire summary data set 'dd.sum'.)
```{r warning = FALSE, message = FALSE}
# renderLeaflet({

# What to show...?
dd.s <- topn.Duration

# Create an indicator for grouping.
groups = as.character(unique(dd.s$KeyPlus))

# The basic map.
map = leaflet(dd.s) %>% 
  # addProviderTiles(providers$CartoDB.Positron)
  # addProviderTiles(providers$Esri.WorldTerrain)
  addProviderTiles(providers$CartoDB.DarkMatter)

# Colors of a specific palette assigned to 'Maximum.Wind'.
groupColors = colorNumeric(palette = "YlOrRd", domain = dd$MaxWindKPH)
# groupColors = scale_colour_brewer("dd.s$Saffir.Simpson")

# Grouping!
for (g in groups) {
  d = dd[dd$KeyPlus == g, ]
  
  map = map %>% 
    addPolylines(data = d,
                 color = "#788E95",
                 group = g,
                 lng = ~ Lon,
                 lat = ~ Lat,
                 weight = 0.6,
                 opacity = 0.6) %>%
    addCircleMarkers(data = d,
                     group = g,
                     lng = ~Lon, 
                     lat = ~Lat, 
                     color = ~groupColors(WindKPH),
                     # color = ~groupColors(Saffir.Simpson),
                     weight = 2,
                     # fill = FALSE,
                     radius = ~(WindKPH^1.2)/50,
                     # radius = ~sqrt(Maximum.Wind)*2,
                     popup = paste("Name: ", d$Name, "<br>",
                                   "Date: ", date(d$DateTime), "<br>",
                                   "Time: ", strftime(d$DateTime, format="%H:%M:%S",tz = "UTC"), "<br>",
                                   "Status: ", d$Status, "<br>",
                                   "Minimum in central Pressure: ", d$Pressure, "<br>",
                                   "Maximum wind speed: ", d$WindKPH, "km/h"))
}

map %>%
  addLayersControl(overlayGroups = groups)
```


Now, which are the storms that showed the highest wind speeds?
```{r warning = FALSE, message = FALSE}
# What to show...?
dd.s <- topn.WindKPH.Max

# Create an indicator for grouping.
groups = as.character(unique(dd.s$KeyPlus))

# The basic map.
map = leaflet(dd.s) %>% 
  # addProviderTiles(providers$CartoDB.Positron)
  # addProviderTiles(providers$Esri.WorldTerrain)
  addProviderTiles(providers$CartoDB.DarkMatter)

# Colors of a specific palette assigned to 'Maximum.Wind'.
groupColors = colorNumeric(palette = "YlOrRd", domain = dd$MaxWindKPH)
# groupColors = scale_colour_brewer("dd.s$Saffir.Simpson")

# Grouping!
for (g in groups) {
  d = dd[dd$KeyPlus == g, ]
  
  map = map %>% 
    addPolylines(data = d,
                 color = "#788E95",
                 group = g,
                 lng = ~ Lon,
                 lat = ~ Lat,
                 weight = 0.6,
                 opacity = 0.6) %>%
    addCircleMarkers(data = d,
                     group = g,
                     lng = ~Lon, 
                     lat = ~Lat, 
                     color = ~groupColors(WindKPH),
                     weight = 2,
                     radius = ~(WindKPH^1.2)/50,   # pronouncing differences
                     popup = paste("Name: ", d$Name, "<br>",
                                   "Date: ", date(d$DateTime), "<br>",
                                   "Time: ", strftime(d$DateTime, format="%H:%M:%S",tz = "UTC"), "<br>",
                                   "Status: ", d$Status, "<br>",
                                   "Minimum in central Pressure: ", d$Pressure, "<br>",
                                   "Maximum wind speed: ", d$WindKPH, "km/h"))
}

map %>%
  addLayersControl(overlayGroups = groups)
```





