---
layout: page
title: ""
date: 2018-01-14
author: "Thomas M. Massie"
output:
  html_document:
    keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include = FALSE}
rm(list=ls())
```


```{r include = FALSE}
# Load libraries.
library(repmis)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(lubridate)
library(scales)
library(chron)
library(ggmap)
library(sp)
library(leaflet)
library(maps)
library(maptools)
library(cowplot)
library(RColorBrewer)
library(ggpubr)
library(showtext)
```



# Storm data exploration (1/2)



## Loading the environment

Loading the workspace that has been saved in the first part of the data exploration.
```{r message = FALSE}
# Entire workspace.
source_data("https://github.com/thomassie/Storms/blob/master/Preparation/StormDataWorkSpace.RData?raw=true")
```


This time I will have a look at specific parts of the data set according to the choices I made before (i.e., a defined time interval; see first part of data exploration).
```{r warning = FALSE, echo = FALSE, include = FALSE}
# All storms between 'year.min' and 'year.max'. 
dd.select <- dd %>% 
  filter(year.min <= year(DateTime) & year(DateTime) <= year.max)

# The summary statistics for the selected data.
dd.select.sum <- dd.sum %>%
  filter(Key %in% dd.select$Key)

# Longest duration (selection).
topn.select.Duration      <- arrange(dd.select.sum, desc(dd.select.sum$MaxDuration))[1:n.select,]
# Strongest wind recorded (selection)
topn.select.Strength.Max  <- arrange(dd.select.sum, desc(dd.select.sum$MaxWindKPH))[1:n.select,]
# Lowest pressure (selection)
topn.select.Lowest.       <- arrange(dd.select.sum, -desc(dd.select.sum$MaxWindKPH))[1:n.select,]
# Minimum pressure (selection)
topn.select.Pressure.Min  <- arrange(dd.select.sum, -desc(dd.select.sum$MinPressure))[1:n.select,]
```


Let's have a look at these storms. First, I want to see which are the storms that lasted the longest and where they ocurred. (I use a data set called 'dd.s' indicating a selection of the entire summary data set 'dd.sum'.)
```{r warning = FALSE, message = FALSE}
# What to show...?
dd.s <- topn.Duration

# Create an indicator for grouping.
groups = as.character(unique(dd.s$KeyPlus))

# The basic map.
map = leaflet(dd.s) %>% 
  # addProviderTiles(providers$CartoDB.Positron)
  # addProviderTiles(providers$Esri.WorldTerrain)
  addProviderTiles(providers$CartoDB.DarkMatter)

# Colors of a specific palette assigned to 'Maximum.Wind'.
groupColors = colorNumeric(palette = "YlOrRd", domain = dd$MaxWindKPH)
# groupColors = scale_colour_brewer("dd.s$Saffir.Simpson")

# Grouping!
for (g in groups) {
  d = dd[dd$KeyPlus == g, ]
  
  map = map %>% 
    addPolylines(data = d,
                 color = "#788E95",
                 group = g,
                 lng = ~ Lon,
                 lat = ~ Lat,
                 weight = 0.6,
                 opacity = 0.6) %>%
    addCircleMarkers(data = d,
                     group = g,
                     lng = ~Lon, 
                     lat = ~Lat, 
                     color = ~groupColors(WindKPH),
                     # color = ~groupColors(Saffir.Simpson),
                     weight = 2,
                     # fill = FALSE,
                     radius = ~(WindKPH^1.2)/50,
                     # radius = ~sqrt(Maximum.Wind)*2,
                     popup = paste("Name: ", d$Name, "<br>",
                                   "Date: ", date(d$DateTime), "<br>",
                                   "Time: ", strftime(d$DateTime, format="%H:%M:%S",tz = "UTC"), "<br>",
                                   "Status: ", d$Status, "<br>",
                                   "Minimum in central Pressure: ", d$Pressure, "<br>",
                                   "Maximum wind speed: ", d$WindKPH, "km/h"))
}

map %>%
  addLayersControl(overlayGroups = groups)
```


Now, which are the storms that showed the highest wind speeds?
```{r warning = FALSE, message = FALSE}
# What to show...?
dd.s <- topn.WindKPH.Max

# Create an indicator for grouping.
groups = as.character(unique(dd.s$KeyPlus))

# The basic map.
map = leaflet(dd.s) %>% 
  # addProviderTiles(providers$CartoDB.Positron)
  # addProviderTiles(providers$Esri.WorldTerrain)
  addProviderTiles(providers$CartoDB.DarkMatter)

# Colors of a specific palette assigned to 'Maximum.Wind'.
groupColors = colorNumeric(palette = "YlOrRd", domain = dd$MaxWindKPH)
# groupColors = scale_colour_brewer("dd.s$Saffir.Simpson")

# Grouping!
for (g in groups) {
  d = dd[dd$KeyPlus == g, ]
  
  map = map %>% 
    addPolylines(data = d,
                 color = "#788E95",
                 group = g,
                 lng = ~ Lon,
                 lat = ~ Lat,
                 weight = 0.6,
                 opacity = 0.6) %>%
    addCircleMarkers(data = d,
                     group = g,
                     lng = ~Lon, 
                     lat = ~Lat, 
                     color = ~groupColors(WindKPH),
                     # color = ~groupColors(Saffir.Simpson),
                     weight = 2,
                     # fill = FALSE,
                     # radius = ~(Pressure^1.2)/500,
                     radius = ~(WindKPH^1.2)/50,
                     # radius = ~sqrt(Maximum.Wind)*2,
                     popup = paste("Name: ", d$Name, "<br>",
                                   "Date: ", date(d$DateTime), "<br>",
                                   "Time: ", strftime(d$DateTime, format="%H:%M:%S",tz = "UTC"), "<br>",
                                   "Status: ", d$Status, "<br>",
                                   "Minimum in central Pressure: ", d$Pressure, "<br>",
                                   "Maximum wind speed: ", d$WindKPH, "km/h"))
}

map %>%
  addLayersControl(overlayGroups = groups)
```


It seems that central pressure and wind speed are correlated with each other. How much? I will check for a linear relationship.
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
rho.wind.pres <- cor(dd$Pressure, dd$WindKPH, use = "complete.obs")
# ggplot(data = dd.345,
#        aes(x = DateTime,
#            y = Pressure)) +
#   # geom_hline(yintercept = 209) +
#   geom_point(aes(size = (WindKPH^1.2)/50,
#                  colour = ???),
#              alpha = 0.3) +
#   scale_colour_manual(values = c("#3F5AA2", "#A4C95D")) +
#   theme_bw()

ggplot(data = dd,
       aes(x = Pressure,
           y = WindKPH)) +
  # geom_hline(yintercept = 209) +
  geom_point(aes(colour = Ocean),
             alpha = 0.4,
             size = 1.1) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  geom_smooth(method = "lm",
              color = "#333333") +
  # geom_smooth(aes(group = Ocean,
  #                 colour = Ocean),
  #             size = 0.5,
  #             method = "lm") +
  geom_hline(yintercept = 0, size = 0.8, colour = "#3C3C3C") +
  labs(x = "Minimum Pressure (hPa)", 
       y = "Maximum wind speed (km/h)",
       title = "Strong relationship between wind speed and \ncentral pressure",
       subtitle = paste( "Recorded maximum wind speed is correlated with central pressure by", expression(rho), " = ", paste(round(rho.wind.pres, 2)),"."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = -0.5, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0, vjust = -2, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1, vjust = -0.2, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.title = element_blank(),
        legend.justification=c(0,1),
        legend.position=c(0.8, 0.97),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(size = 10, colour = "#3C3C3C")) +
  # annotate("text", x = 920, y = 310, label = paste("rho = ", round(rho.wind.pres, 2))) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```


Not bad, but the residuals (data - fit) seem to show some [heteroskedasticity](https://en.wikipedia.org/wiki/Heteroscedasticity).  Maybe the relationship can be better described by another fit...?
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
rho.wind.pres <- cor(dd$Pressure, dd$WindKPH, use = "complete.obs")
# ggplot(data = dd.345,
#        aes(x = DateTime,
#            y = Pressure)) +
#   # geom_hline(yintercept = 209) +
#   geom_point(aes(size = (WindKPH^1.2)/50,
#                  colour = ???),
#              alpha = 0.3) +
#   scale_colour_manual(values = c("#3F5AA2", "#A4C95D")) +
#   theme_bw()

ggplot(data = dd,
       aes(x = Pressure,
           y = WindKPH)) +
  # geom_hline(yintercept = 209) +
  geom_point(aes(colour = Ocean),
             alpha = 0.4,
             size = 1.1) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  # geom_smooth(formula = y ~ -x^2,
  #             color = "#333333") +
  geom_smooth(method = "lm", 
              formula = y ~ splines::bs(x, 3),
              color = "#333333") +
  # geom_smooth(aes(group = Ocean,
  #                 colour = Ocean),
  #             size = 0.5,
  #             method = "lm") +
  geom_hline(yintercept = 0, size = 0.8, colour = "#3C3C3C") +
  labs(x = "Minimum Pressure (hPa)", 
       y = "Maximum wind speed (km/h)",
       title = "Strong relationship between wind speed and \ncentral pressure",
       subtitle = paste("A non-linear fit represents this relationship a bit better."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = -0.5, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0, vjust = -2, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1, vjust = -0.2, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.title = element_blank(),
        legend.justification=c(0,1),
        legend.position=c(0.8, 0.95),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(size = 10, colour = "#3C3C3C")) +
  # annotate("text", x = 920, y = 310, label = paste("rho = ", round(rho.wind.pres, 2))) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```


When do the storms typicall occur during a year?
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.pressure.year.main <- ggplot(data = dd,
                                  aes(x = DateTimeSameYear,
                                      y = Pressure)) + 
  geom_point(aes(colour = Ocean),
             size = 0.8,
             alpha = 0.1) +
  # geom_smooth(method = "lm",
  #             color = "#333333") +
  labs(x = "Month", 
       y = "Minimum central pressure (mbar)",
       title = "When do hurricans and typhoons occur?",
       subtitle = expression("Annual minima in central pressure records."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = -0.5, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0, vjust = -5, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1.6, vjust = -0.1, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.title = element_blank(),
        legend.justification=c(0,1),
        legend.position=c(1.02, 0.3),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(size = 10)) +
  # geom_text(data = dd.lines, aes(label = year(int), x = int, y = -Inf), angle = 0, inherit.aes = F, hjust = -0.2, vjust = -36, size = 3.5) +
  # ggpubr::color_palette("jco") +
  # scale_colour_manual(values = c("#B78A3F", "#58758C")) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  # rremove("legend") +
  scale_x_datetime(labels = date_format("%b"),
                   date_breaks = "month",
                   limits = c(min(dd$DateTimeSameYear), max(dd$DateTimeSameYear))) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

# A density plot on top of the main plot.
# TREAT the date.time AS NUMERIC!!!
plot.pressure.year.dens.x <- axis_canvas(plot.pressure.year.main, axis = "x") + 
  geom_density(data = dd, aes(x = as.numeric(DateTimeSameYear), fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # ggpubr::color_palette("jco")
  # scale_fill_manual(values = c("#B78A3F", "#58758C"))
  scale_fill_manual(values = c("#FF281E", "#0090CF"))

# ...and one on the right site.
plot.pressure.year.dens.y <- axis_canvas(plot.pressure.year.main, axis = "y", coord_flip = TRUE) +
  geom_density(data = dd, aes(x = as.numeric(Pressure), fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # ggpubr::color_palette("jco") +
  # scale_fill_manual(values = c("#B78A3F", "#58758C")) +
  scale_fill_manual(values = c("#FF281E", "#0090CF")) +
  coord_flip()
```

Combined to one figure...
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.pressure.year.1 <- insert_xaxis_grob(plot.pressure.year.main, 
                                          plot.pressure.year.dens.x, 
                                          grid::unit(0.2, "null"), 
                                          position = "top")
plot.pressure.year.2 <- insert_yaxis_grob(plot.pressure.year.1, 
                                          plot.pressure.year.dens.y, 
                                          grid::unit(.2, "null"), 
                                          position = "right")
ggdraw(plot.pressure.year.2)
```


When do the storms typicall occur during a year?
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.duration.main <- ggplot(data = dd,
                             aes(x = Duration,
                                 y = Pressure)) + 
  geom_vline(xintercept = 0, size = 0.8, colour = "#3C3C3C") +
  geom_point(aes(colour = Ocean),
             size = 0.8,
             alpha = 0.1) +
  # geom_point(data = dd.sum,
  #            aes(x = MinPressureAfter,
  #                y = MinPressure),
  #            colour = "#666666",
  #            size = 0.8,
  #            alpha = 0.4) +
  # geom_smooth(method = "lm",
  #             color = "#333333") +
  labs(x = "Duration (days)", 
       y = "Minimum central pressure (mbar)",
       title = expression("When do storms reach their maximum intensity?"),
       subtitle = expression("Most storms showed their lowest minimum in \ncentral pressure briefly after they built up."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = -0.5, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0.9, vjust = -7, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1.6, vjust = -0.1, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.title = element_blank(),
        legend.justification=c(0,1),
        legend.position=c(1.02, 0.3),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(size = 10)) +
  # geom_text(data = dd.lines, aes(label = year(int), x = int, y = -Inf), angle = 0, inherit.aes = F, hjust = -0.2, vjust = -36, size = 3.5) +
  # ggpubr::color_palette("jco") +
  # scale_colour_manual(values = c("#B78A3F", "#58758C")) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  # rremove("legend") +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

# A density plot on top of the main plot.
# TREAT the date.time AS NUMERIC!!!
plot.duration.dens.x <- axis_canvas(plot.duration.main, axis = "x") + 
  geom_density(data = dd, aes(x = Duration, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # ggpubr::color_palette("jco")
  # scale_fill_manual(values = c("#B78A3F", "#58758C"))
  scale_fill_manual(values = c("#FF281E", "#0090CF"))

# ...and one on the right site.
plot.duration.dens.y <- axis_canvas(plot.duration.main, axis = "y", coord_flip = TRUE) +
  geom_density(data = dd, aes(x = Pressure, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # ggpubr::color_palette("jco") +
  # scale_fill_manual(values = c("#B78A3F", "#58758C")) +
  scale_fill_manual(values = c("#FF281E", "#0090CF")) +
  coord_flip()
```

Now, one can combine all three plots to a single figure.
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.duration.1 <- insert_xaxis_grob(plot.duration.main, 
                                     plot.duration.dens.x, 
                                     grid::unit(0.2, "null"), 
                                     position = "top")
plot.duration.2 <- insert_yaxis_grob(plot.duration.1, 
                                     plot.duration.dens.y, 
                                     grid::unit(.2, "null"), 
                                     position = "right")
ggdraw(plot.duration.2)
```


After what time do storms typically reach their maximum intensity?
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.duration.storm.main <- ggplot(data = dd.sum,
                                   aes(x = MinPressureAfter,
                                       y = MinPressure)) + 
  geom_vline(xintercept = 0, size = 0.8, colour = "#3C3C3C") +
  geom_point(aes(colour = Ocean,
                 size = MaxWindKPH),
             # size = MaxWindKPH^4/100000),
             show.legend = TRUE,
             # size = 0.8,
             alpha = 0.2) +
  # geom_smooth(method = "lm",
  #             color = "#333333") +
  labs(x = "Duration at lowest minimum pressure (days)", 
       y = "Minimum central pressure (mbar)",
       title = expression("When do storms reach their maximum intensity?"),
       subtitle = expression("Most storms show their lowest minimum \npressure briefly after they built up."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = -0.5, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0.9, vjust = -7, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1.6, vjust = -0.1, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.justification=c(0,1),
        # legend.title = element_blank(),
        legend.position=c(1.02, 0.3),
        legend.background = element_blank(),
        legend.title=element_text(size=10),
        # legend.key = element_blank(),
        legend.text = element_text(size = 10)) +
  labs(colour = "Ocean",
       size = expression("Wind speed \n(km/h)")) +
  # geom_text(data = dd.lines, aes(label = year(int), x = int, y = -Inf), angle = 0, inherit.aes = F, hjust = -0.2, vjust = -36, size = 3.5) +
  # ggpubr::color_palette("jco") +
  # scale_colour_manual(values = c("#B78A3F", "#58758C")) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  # rremove("legend") +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

# A density plot on top of the main plot.
# TREAT the date.time AS NUMERIC!!!
plot.duration.storm.dens.x <- axis_canvas(plot.duration.storm.main, axis = "x") + 
  geom_density(data = dd.sum, aes(x = MinPressureAfter, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # ggpubr::color_palette("jco")
  # scale_fill_manual(values = c("#B78A3F", "#58758C"))
  scale_fill_manual(values = c("#FF281E", "#0090CF"))

# ...and one on the right site.
plot.duration.storm.dens.y <- axis_canvas(plot.duration.storm.main, axis = "y", coord_flip = TRUE) +
  geom_density(data = dd.sum, aes(x = MinPressure, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # ggpubr::color_palette("jco") +
  # scale_fill_manual(values = c("#B78A3F", "#58758C")) +
  scale_fill_manual(values = c("#FF281E", "#0090CF")) +
  coord_flip()
```

Now, one can combine all three plots to a single figure.
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.duration.storm.1 <- insert_xaxis_grob(plot.duration.storm.main, 
                                           plot.duration.storm.dens.x, 
                                           grid::unit(0.2, "null"), 
                                           position = "top")
plot.duration.storm.2 <- insert_yaxis_grob(plot.duration.storm.1, 
                                           plot.duration.storm.dens.y, 
                                           grid::unit(.2, "null"), 
                                           position = "right")
ggdraw(plot.duration.storm.2)
```


However, the time at which a storm reaches its maximum intensity has to be related to its total duration.
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.duration.storm.main <- ggplot(data = dd.sum,
                                   aes(x = MinPressureAfter/MaxDuration *100,
                                       y = MinPressure)) + 
  geom_vline(xintercept = 0, size = 0.8, colour = "#3C3C3C") +
  geom_point(aes(colour = Ocean,
                 size = MaxWindKPH),
             # size = MaxWindKPH^4/100000),
             show.legend = TRUE,
             # size = 0.8,
             alpha = 0.2) +
  # geom_smooth(method = "lm",
  #             color = "#333333") +
  labs(x = "Duration at lowest minimum pressure in relation to total duration (%)", 
       y = "Minimum central pressure (mbar)",
       title = expression("When do storms reach their maximum intensity?"),
       subtitle = expression("Most storms are at their maximum intensity at around half-time."),
       caption = "Source: NOAA's National Hurricane Center (http://www.nhc.noaa.gov/data/)") +
  theme(axis.text = element_text(family = "Varela Round"),
        axis.text.x = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 1),
        axis.text.y = element_text(size = 11, colour = "#3C3C3C", face = "bold", vjust = 0.5),
        axis.ticks = element_line(colour = "#D7D8D8", size = 0.2),
        axis.ticks.length = unit(5, "mm"),
        axis.line = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0, vjust = 0, colour = "#3C3C3C", size = 20),
        plot.subtitle = element_text(hjust = 0, vjust = 0.5, colour = "#3C3C3C", size = 11),
        plot.caption = element_text(size = 8, hjust = 1.6, vjust = -0.1, colour = "#7F8182"),
        panel.background = element_rect(fill = "#FAFAF2"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "#FAFAF2", colour = "#FAFAF2"),
        panel.grid.major = element_line(colour = "#D7D8D8", size = 0.2),
        panel.grid.minor = element_line(colour = "#D7D8D8", size = 0.2)) +
  theme(legend.justification=c(0,1),
        # legend.title = element_blank(),
        legend.position=c(1.02, 0.3),
        legend.background = element_blank(),
        legend.title=element_text(size=10),
        # legend.key = element_blank(),
        legend.text = element_text(size = 10)) +
  labs(colour = "Ocean",
       size = expression("Wind speed \n(km/h)")) +
  # geom_text(data = dd.lines, aes(label = year(int), x = int, y = -Inf), angle = 0, inherit.aes = F, hjust = -0.2, vjust = -36, size = 3.5) +
  # ggpubr::color_palette("jco") +
  # scale_colour_manual(values = c("#B78A3F", "#58758C")) +
  scale_colour_manual(values = c("#FF281E", "#0090CF")) +
  # rremove("legend") +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

# A density plot on top of the main plot.
# TREAT the date.time AS NUMERIC!!!
plot.duration.storm.dens.x <- axis_canvas(plot.duration.storm.main, axis = "x") + 
  geom_density(data = dd.sum, aes(x = MinPressureAfter/MaxDuration *100, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # geom_histogram(data = dd.sum, aes(x = MinPressureAfter/MaxDuration *100, fill = Ocean),
  #                alpha = 0.6, size = 0.2,
  #                binwidth = 2) +
  # ggpubr::color_palette("jco")
  # scale_fill_manual(values = c("#B78A3F", "#58758C"))
  scale_fill_manual(values = c("#FF281E", "#0090CF"))

# ...and one on the right site.
plot.duration.storm.dens.y <- axis_canvas(plot.duration.storm.main, axis = "y", coord_flip = TRUE) +
  geom_density(data = dd.sum, aes(x = MinPressure, fill = Ocean),
               alpha = 0.6, size = 0.2) +
  # geom_histogram(data = dd.sum, aes(x = MinPressure, fill = Ocean),
  #                alpha = 0.6, size = 0.2,
  #                binwidth = 2) +
  # ggpubr::color_palette("jco") +
  # scale_fill_manual(values = c("#B78A3F", "#58758C")) +
  scale_fill_manual(values = c("#FF281E", "#0090CF")) +
  coord_flip()
```

Now, one can combine all three plots to a single figure.
```{r fig.height = 6, fig.width = 8, messages = FALSE, warning = FALSE}
plot.duration.storm.1 <- insert_xaxis_grob(plot.duration.storm.main, 
                                           plot.duration.storm.dens.x, 
                                           grid::unit(0.2, "null"), 
                                           position = "top")
plot.duration.storm.2 <- insert_yaxis_grob(plot.duration.storm.1, 
                                           plot.duration.storm.dens.y, 
                                           grid::unit(.2, "null"), 
                                           position = "right")
ggdraw(plot.duration.storm.2)
```


Next, one can ask for the regions that are most storm prone.
```{r}
plot.pressure.main <- ggplot(data = dd,
                             aes(x = Lon -360,
                                 y = Lat)) +
  stat_density_2d(geom = "raster",
                  aes(fill = ..density..),
                  contour = FALSE,
                  alpha = 0.9,
                  show.legend = FALSE)
# scale_fill_gradientn(colours = rev( brewer.pal( 9, "Blues" ))) +
```


















