---
title: "Storm Data Analysis"
author: "Thomas M. Massie"
date: "10 1 2018"
output: 
  html_document: 
    keep_md: true
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include = FALSE}
rm(list=ls())
```


```{r include = FALSE}
# Load libraries.
library(repmis)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(lubridate)
library(scales)
library(chron)
library(ggmap)
library(sp)
library(leaflet)
library(maps)
library(maptools)
library(RColorBrewer)
```


# Storm data analysis



## Loading the dataset

```{r}
# Entire workspace.
source_data("https://github.com/thomassie/Storms/blob/master/Data/StormDataWS.RData?raw=true")
```



## Exploring the dataset

First, I generate two dataset that include a couple of summary statistics such as minimum pressure or maximum duration for each **storm**...
```{r}
# Some summary statistics to see which was the strongest storm etc.
dd.sum <- dd %>%
  group_by(ID) %>%
  summarise(Name = unique(Name),
            ID.plus = unique(ID.plus),
            Year = first(Year),
            Min.Pressure = min(Minimum.Pressure, na.rm = TRUE),
            Max.Duration = max(Duration, na.rm = TRUE),
            Mean.Strength.kph = round(mean(Maximum.Wind.kph, 0)),
            Max.Strength.kph = max(Maximum.Wind.kph, na.rm = TRUE),
            Storm.Start = format(DateTime.new[1], "%d/%m %H:%M:%S"),
            Storm.Stop = format(DateTime.new[length(DateTime.new)], "%d/%m %H:%M:%S")) %>%
  arrange(-desc(Year))
str(dd.sum)
```

...and for each **year**. I'll use this dataset mainly for mapping. You'll see this at a later point.
```{r}
dd.sum.year <- dd %>%
  group_by(Year) %>%
  summarise(DateTime.new = first(DateTime.new),
            # Date = first(DateTime.new[which(Minimum.Pressure == min(Minimum.Pressure, na.rm = TRUE))]),
            Min.Pressure = min(Minimum.Pressure, na.rm = TRUE),
            Max.Duration = max(Duration, na.rm = TRUE),
            Mean.Strength.kph = round(mean(Maximum.Wind.kph, 0)),
            Max.Strength.kph = max(Maximum.Wind.kph, na.rm = TRUE)) %>%
  arrange(-desc(Year))
```


But first, I have to assign the exact dates and times for these annual extreme values.
```{r}
# Get exact date and time for all extreme values!
p.min <- rep(NA, length(unique(dd$Year)))
d.max <- p.min
w.max <- p.min
for (i in 1:length(unique(dd$Year))) {
  p.min[i] = first(filter(dd, 
                          Year == unique(dd$Year)[i] & 
                            Minimum.Pressure == as.character(dd.sum.year$Min.Pressure[i]))$DateTime.new)
  d.max[i] = first(filter(dd, 
                          Year == unique(dd$Year)[i] & 
                            Duration == as.character(dd.sum.year$Max.Duration[i]))$DateTime.new)
  w.max[i] = first(filter(dd, 
                          Year == unique(dd$Year)[i] & 
                            Maximum.Wind.kph == as.character(dd.sum.year$Max.Strength.kph[i]))$DateTime.new)
}

dd.sum.year <- dd.sum.year %>%
  mutate(DateTime.p = as.POSIXct(p.min, origin = "1970-01-01")) %>%
  mutate(DateTime.d = as.POSIXct(d.max, origin = "1970-01-01")) %>%
  mutate(DateTime.w = as.POSIXct(w.max, origin = "1970-01-01"))

str(dd.sum.year)
```



Then I make a couple of choices to get a nice ranking for a specific time period I would like to put my focus on.
```{r}
# The first 'n.select' storms are selected for ranking
n.select <- 10
# Time window: from 'year.min' to 'year.max'. 
# Fill in identical years to get data only of year x.
year.min <- 2012
year.max <- 2015
```


A first glimpse at the rankings of the **entire** data.
```{r}
# Ranking according to...
# ...longest duration (enire data set).
topn.Duration       <- arrange(dd.sum, desc(dd.sum$Max.Duration))[1:n.select,] %>%
  mutate(criterion = "duration")
# Rearrange levels.
topn.Duration$ID.plus <- factor(topn.Duration$ID.plus, 
                                levels = topn.Duration$ID.plus[desc(topn.Duration$Max.Duration)])
# ...strongest wind speed recorded (entire data set)
topn.Strength.Max  <- arrange(dd.sum, desc(dd.sum$Max.Strength.kph))[1:n.select,] %>%
  mutate(criterion = "wind")
# Rearrange levels.
topn.Strength.Max$ID.plus <- factor(topn.Strength.Max$ID.plus, 
                                levels = topn.Strength.Max$ID.plus[desc(topn.Strength.Max$Max.Strength.kph)])
# ...minimum pressure (entire data set)
topn.Pressure.Min  <- arrange(dd.sum, -desc(dd.sum$Min.Pressure))[1:n.select,] %>%
  mutate(criterion = "pressure")
# Rearrange levels.
topn.Pressure.Min$ID.plus <- factor(topn.Pressure.Min$ID.plus, 
                                levels = topn.Pressure.Min$ID.plus[desc(topn.Pressure.Min$Min.Pressure)])
#dd.topn <- rbind(topn.Duration, topn.Strength.Max, topn.Pressure.Min)
```


```{r}
topn.Duration %>%
  ggplot(aes(x = ID.plus,
             y = Max.Duration)) +
  geom_bar(stat="identity") +
  geom_col() +
  # theme_bw() +
  theme_classic() +
  xlab("") +
  ylab("") 
+
  theme(axis.text = element_text(size = 50),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank()) +
  geom_text(aes(label = frequency),
            size = 14,
            color = "white",
            position = position_dodge(width = 0.9),
            hjust = 1.2) +
  theme(legend.position="none")
  
scale_fill_manual(values = colorNumeric(palette = "YlOrRd", domain = topn.Duration$Max.Duration)) +
  theme_()





library("cowplot")
plot_grid(bxp, dp, bp + rremove("x.text"), 
          labels = c("A", "B", "C"),
          ncol = 2, nrow = 2)
```



Now I select specific parts of the prepared dataset accrding to the choices I made.
```{r}
# All storms between 'year.min' and 'year.max'. 
dd.select <- dd %>% 
  filter(year.min <= year(Date.new) & year(Date.new) <= year.max)

# The summary statistics for the selected data.
dd.select.sum <- dd.sum %>%
  filter(ID %in% dd.select$ID)

# Longest duration (selection).
topn.select.Duration      <- arrange(dd.select.sum, desc(dd.select.sum$Max.Duration))[1:n.select,]
# Strongest wind recorded (selection)
topn.select.Strength.Max  <- arrange(dd.select.sum, desc(dd.select.sum$Max.Strength.kph))[1:n.select,]
# Lowest pressure (selection)
topn.select.Lowest.       <- arrange(dd.select.sum, -desc(dd.select.sum$Max.Strength.kph))[1:n.select,]
# Minimum pressure (selection)
topn.select.Pressure.Min  <- arrange(dd.select.sum, -desc(dd.select.sum$Min.Pressure))[1:n.select,]
```


Let's have a look at these storms. First, I want to see which are the storms that lasted the longest. (I use a dataset called 'dd.s' indicating a selection of the entire summary data set 'dd.sum'.)
```{r}
dd.s <- topn.Duration

# Create an indicator for grouping.
groups = as.character(unique(dd.s$ID.plus))

# The basic map.
map = leaflet(dd.s) %>% 
  # addProviderTiles(providers$CartoDB.Positron)
  # addProviderTiles(providers$Esri.WorldTerrain)
  addProviderTiles(providers$CartoDB.DarkMatter)

# Colors of a specific palette assigned to 'Maximum.Wind'.
groupColors = colorNumeric(palette = "YlOrRd", domain = dd.s$Maximum.Wind.kph)

# Grouping!
for (g in groups) {
  d = dd[dd$ID.plus == g, ]
  map = map %>% 
    addPolylines(data = d,
                 color = "#788E95",
                 group = g,
                 lng = ~ Longitude,
                 lat = ~ Latitude,
                 weight = 0.6,
                 opacity = 0.6) %>%
    addCircleMarkers(data = d,
                     group = g,
                     lng = ~Longitude, 
                     lat = ~Latitude, 
                     color = ~groupColors(Maximum.Wind.kph),
                     weight = 2,
                     # fill = FALSE,
                     radius = ~(Maximum.Wind^1.2)/50,
                     # radius = ~sqrt(Maximum.Wind)*2,
                     popup = paste("Name: ", d$Name, "<br>",
                                   "Date: ", d$Date.new, "<br>",
                                   "Time: ", d$Time.new, "<br>",
                                   "Status: ", d$Status, "<br>",
                                   "Maximum wind speed: ", d$Maximum.Wind.kph, "km/h", "<br>"))
}

map %>%
  addLayersControl(overlayGroups = groups)
```


```{r}
dd.s %>%
  ggplot() +
  aes(x = DateTime.new,
      y = Minimum.Pressure
  ) +
  stat_density_2d(geom = "raster", 
                  aes(fill = ..density..), 
                  contour = FALSE,
                  alpha = 0.9,
                  show.legend = FALSE) +
  scale_fill_gradientn(colours = rev( brewer.pal( 9, "Greys" ))) +
  geom_point(alpha = 0.5,
             size = 0.3,
             colour = "#FBFBFA") +
  geom_smooth(method = "lm",
              se = FALSE,
              span = 0.2,
              colour = "#F93A2E",
              alpha = 0.9,
              size = 1.2) +
  geom_smooth(data = filter(dd.sum.year, year.min <= year(dd.sum.year$DateTime.p) & year(dd.sum.year$DateTime.p) <= year.max),
              aes(x = DateTime.p,
                  y = Min.Pressure),
              method = "lm",
              se = TRUE,
              span = 0.2,
              linetype = "dashed",
              colour = "#F93A2E",
              alpha = 0.9,
              size = 0.5) +
  geom_point(data = filter(dd.sum.year, year.min <= year(dd.sum.year$DateTime.p) & year(dd.sum.year$DateTime.p) <= year.max),
             aes(x = DateTime.p,
                 y = Min.Pressure),
             shape = 1,
             size = 1.2,
             stroke = 0.8,
             colour = "#F93A2E") +
  geom_line(data = filter(dd.sum.year, year.min <= year(dd.sum.year$DateTime.new) & year(dd.sum.year$DateTime.p) <= year.max),
            aes(x = (DateTime.p),
                y = Min.Pressure),
            size = 0.3,
            alpha = 0.9,
            colour = "#F93A2E") +
  theme_bw() +
  geom_rug(alpha = 0.025) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(x = "Year", y = "Minimum Pressure (mbar)",
       title = "")
```











